Exploring Alternative Routes Using Multipath TCP

### 2017-06-04 paper/presentation.tex
presentation: switch to bytefield where I can
presentation: supplement Intro w/ application-layer arguments

\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{wrapfig}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\setbeamertemplate{note page}[plain]
\setbeameroption{show notes on second screen=right}
\setbeamertemplate{navigation symbols}{}%remove navigation symbols
\AtBeginNote{%
    \let\enumerate\itemize%
    \let\endenumerate\enditemize%
}

@@

    \begin{itemize}
    \item Not simply something we can hypothesize about.
    \item Real world examples show that default routes aren't the best paths for
      users.
    \item In fact, many times you can add a ``detour'' to your route and end
      up with an improved path.
    \item At right, the default path from Los Altos to Walnut Creek has round
      trip time of 137 milliseconds, while the constituent paths of a detour
      route through Santa Clara add up to 40 milliseconds
    \end{itemize}

@@

  \begin{itemize}
  \item Not just isolated examples.
  \item Same study looked at a group of computers across universities
  \item Observed pairwise RTT, loss rate.
  \item Searched for pairs where ``triangle inequality violated'', i.e. lower
    RTT or loss rate via detour.
  \item In graph, blue is default route, yellow is best alternative route, and
    we compare loss rate
  \item For majority of pairs, alternative route is more reliable.
  \item For some pairs, reliability difference is staggering
  \end{itemize}

@@

\begin{frame}
  \frametitle{Access Link Underutilization}
  % this slide reads as very stupid - at least the first bullet point does.
  % let's think about making a better case for improving links, network core
  % bottlenecks, etc

  \begin{itemize}
  \item Fiber-to-the-home is coming!
  \item However, residential bandwidth is not fully utilized
    \mycite{fibertothehome}
    \begin{itemize}
    \item Short-lived TCP sessions?
    \item Small TCP initial window?
    \item Network core can't support bandwidth?
    \end{itemize}
  \item Using alternative routes can improve performance
  \item Aggregating multiple routes can perform even better
  \end{itemize}

@@

@@

  \textbf{Problem:} Internet routing inefficiencies can prevent TCP connections
  from achieving their full potential throughput.

@@

  \textbf{Solution:} A system which can use and aggregate alternative routes for
  TCP connections.

@@

  \begin{itemize}
  \item A mechanism for adding alternative routes to Multipath TCP (MPTCP)
    sessions among single-homed hosts.
  \item An evaluation of this mechanism on emulated and real-world networks.
  \end{itemize}
  \note{

    \begin{itemize}
    \item Putting together, problem: routing inefficiencies can prevent TCP
      connections from achieving throughput.
    \item Furthermore, while detour approaches are plentiful, and even a few
      which use multiple paths, as well as application-level strategies, there
      are no strategies which are usable by unmodified applications
    \item A solution could be a system which uses and aggregates alternative
      routes for MPTCP connections.
    \item Contribution 1 - a mechanism for adding new routes to MPTCP
      connections which are single-homed, with no modification to applications.
    \item Contribution 2 - experimental validation that the mechanism works,
      even at higher throughputs
    \end{itemize}

@@

  \begin{center}
    \vfill
\begin{BVerbatim}
+-----------------------------------+
| ver(1) | op(1)  | reserved (2)    |
+-----------------------------------+
|           rip (4 bytes)           |
+-----------------------------------+
| rpt (2 bytes)   | dpt (2 bytes)   |
+-----------------|-----------------+
\end{BVerbatim}
    \vfill

>>>>

\usepackage{appendixnumberbeamer}
\usepackage{pgfpages}
\usepackage{wrapfig}
\usepackage[export]{adjustbox}
\usepackage{graphicx}
\usepackage{fancyvrb}
\usepackage{bytefield}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\setbeamertemplate{note page}[plain]
\setbeameroption{show notes on second screen=right}
\setbeamertemplate{navigation symbols}{}%remove navigation symbols
\AtBeginNote{%
    \let\enumerate\itemize%
    \let\endenumerate\enditemize%
}

@@

    \begin{itemize}
    \item Not simply something we can hypothesize about.
    \item Real world examples show that default routes aren't the best paths for
      users.
    \item In fact, many times you can add a ``detour'' to your route and end
      up with an improved path.
    \item At right, the default path from Los Altos to Walnut Creek has round
      trip time of 137 milliseconds, while the constituent paths of a detour
      route through Santa Clara add up to 30 milliseconds
    \end{itemize}

@@

  \begin{itemize}
  \item Not just isolated examples.
  \item Same study looked at a group of computers across universities
  \item Observed pairwise RTT, loss rate.
  \item Searched for pairs where ``triangle inequality violated'', i.e. lower
    RTT or loss rate via detour.
  \item In graph, blue is default route, yellow is best alternative route, and
    we compare loss rate
  \item For 80\% of pairs, alternative route is more reliable. For 25\% of
    pairs, alternative route improves loss rate by at least a factor of 6
  \item For some pairs, reliability difference is staggering!
  \end{itemize}

@@

\begin{frame}
  \frametitle{Access Link Underutilization}
  \begin{itemize}
  \item Residential bandwidth constantly improves
  \item However, residential bandwidth is not fully utilized
    \mycite{fibertothehome}
    \begin{itemize}
    \item Short-lived TCP sessions?
    \item Anemic send buffers?
    \item Network core can't support bandwidth?
    \end{itemize}
  \item Using alternative routes can improve performance
  \item Aggregating multiple routes can perform even better
  \end{itemize}

@@

\TitledFigureSlide{figures/Concept.pdf}{Concept}{
  \begin{itemize}
  \item Detour routing has shown promise for a long time
  \item Some applications have been able to leverage it (e.g. Bittorrent)
  \item However, few unmodified applications may use the detour routing strategy
  \item Our concept is to use MPTCP to achieve detour routing in unmodified
    applications
  \item We use tunneling mechanisms to put a subflow across a detour host
  \end{itemize}
}

@@

  \textbf{Problem:} Unmodified applications cannot use detour routing to
  circumvent Internet routing inefficiencies.

@@

  \textbf{Solution:} An OS-level detour routing system that leverages Multipath
  TCP (MPTCP).

@@

  \begin{itemize}
  \item A method for performing detour routing with unmodified applications
  \item A prototype implementation in the Linux kernel
  \item An evaluation of this mechanism on emulated networks and the Internet
  \end{itemize}
  \note{

    \begin{itemize}
    \item Putting together, problem: unmodified applications can't take
      advantage of detour routing solutions to Internet routing inefficiencies.
    \item A solution could be a system which uses and aggregates alternative
      routes for MPTCP connections.
    \item Contribution 1 - a mechanism for adding new routes to MPTCP
      connections which are single-homed, with no modification to applications.
    \item Contribution 2 - experimental validation that the mechanism works,
      even at higher throughputs
    \end{itemize}

@@

  \begin{center}
    \vfill
    \begin{bytefield}[bitwidth=0.6em]{32}
      \bitheader{0,7,8,15,16,23,24,31}\\
      \bitbox{8}{\texttt{ver}} & \bitbox{8}{\texttt{op}} & \bitbox{16}{\texttt{reserved}} \\
      \wordbox{1}{\texttt{rip}} \\
      \bitbox{16}{\texttt{rpt}} & \bitbox{16}{\texttt{dpt}} \\
    \end{bytefield}
    \vfill
